---
title: "BCP HEI"
output: html_notebook
---

```{r setup}
library(plyr)
library(dplyr)
library(utils)
```

# About the data
Intake Properties Totals File (File 04)
Serving Count Totals File (File 09)

####### will redo eventually 
zip <- ".zip"
data_dir_path = "~/Desktop/Baby ConnectoMe Datasets/Concat/"

file_list<-Sys.glob(file.path(paste0(data_dir_path,'*',zip, collapse ='')))
files_zipped<-lapply(X = file_list, FUN = unzip)

fours<-lapply(files_zipped, function(i){
  grep("04",i,value = TRUE)
})

nines<-lapply(files_zipped, function(i){
  grep("09",i,value = TRUE)
})



All4 <- lapply(fours,function(i){
  read.table(i, header=TRUE, sep="\t")
})

All9 <- lapply(nines,function(i){
  read.table(i, header=TRUE, sep="\t")
})

All9[[2]]$Participant.ID
l<-strsplit(as.character(All9[[2]]$Participant.ID), "_")
l
df<-lapply(l, function(i){
  data.frame(matrix(unlist(i), ncol=3, byrow=TRUE))
})

df
df <- data.frame(matrix(unlist(l), byrow=T))
unlist(l)
df
df<-matrix(unlist(strsplit(as.character(All9[[2]]$Participant.ID), "_")), ncol=2, byrow=TRUE)
df




# Important functions
```{r}
adeq_check<-function(inital, new, parameter){
  #parameters are based on the HEI 2015 from NDSR SAS HEI script
  v5<-c(data$hei_totveg, data$hei_greensbeans, data$hei_totfruit, data$hei_wholefruit, data$hei_totproteins, data$hei_seafoodplantprot)
  
  if(inital %in% v5 == TRUE){
    print("neat")
    init= inital/data$energy
    
    new	<-ifelse(init==0, 0, NA)
    new	<-ifelse(init >= parameter, 5, NA)
    new	<-ifelse(is.na(new), 5*(init/parameter),new)
    
  }
}


adeq_check10<-function(inital, new, parameter){
  #parameters are based on the HEI 2015 from NDSR SAS HEI script
  v10<-c(data$hei_wholegrains, data$hei_dairy)
  
  if(inital %in% v10 == TRUE){
    print("cool")
    init= inital/data$energy
    
    new	<-ifelse(init==0, 0, NA)
    new	<-ifelse(init >= parameter, 10, NA)
    new	<-ifelse(is.na(new), 10*(init/parameter),new)
  } 
}


fa_adeq<-function(poly,mono,sat){
  FARMIN=1.2
  FARMAX=2.5
  init=(poly+mono)/sat
  new <-ifelse(init <= FARMIN, 0, NA)
  new <-ifelse(init > FARMAX, 10, NA)
  new	<-ifelse(is.na(new), 10*((init-FARMIN)/(FARMAX-FARMIN)),new)
}

mod_check<-function(input, new, MIN, MAX){
  v1<-c(data$hei_sodium, data$hei_refinedgrains)
  v2<-c(data$hei_addedsugars)
  v3<-c(data$X..Calories.from.SFA)
  if(input %in% v1 == TRUE){
  init=input/data$energy
  
  new <-ifelse(init >= MAX, 0, NA)
  new <-ifelse(init <= MIN, 10, NA)
  new	<-ifelse(is.na(new), 10-(10*((init-MIN)/(MAX-MIN))),new)
  
  }else if(input %in% v2 == TRUE){
    print("sweet")
    init=100*input/data$Energy..kcal.
    
    new <-ifelse(init >= MAX, 0, NA)
    new <-ifelse(init < MIN, 10, NA)
    new	<-ifelse(is.na(new), 10-(10*((init-MIN)/(MAX-MIN))),new)
  }else if(input %in% v3 == TRUE){
    init=input
    
    new <-ifelse(init > MAX, 0, NA)
    new <-ifelse(init < MIN, 10, NA)
    new	<-ifelse(is.na(new), 10-(10*((init-MIN)/(MAX-MIN))),new)
  }
}
```



# Load data
# Notes
Having issues with the text value fields
Need to assess a better way to read in the data 
```{r load_data}
file09<-read.table("~/Google Drive/BCP/data/Concat/Mom_BCP_dataset9.txt", sep="\t", header=T)

file04<-read.table("~/Google Drive/BCP/data/Concat/Mom_BCP_dataset4.txt", sep="\t", header=T)
file04
```

```{r}
data<-join(file04, file09)
dim(data)

# make the data numeric instead of factors
data[,19:392]<-apply(data[,19:392], MARGIN = 2, as.numeric)
```

# Check if the data is non-zero
deciding if I want this
```{r}
data$energy<-data$Energy..kcal./1000 

#ifelse(data$Energy..kcal. > 0, )
#if rikcal=0 then do;
#  HEIX1_TOTALVEG=0; HEIX2_GREEN_AND_BEAN=0; HEIX3_TOTALFRUIT=0; HEIX4_WHOLEFRUIT=0; HEIX5_WHOLEGRAIN=0; HEIX6_TOTALDAIRY=0;
#  HEIX7_TOTPROT=0;  HEIX8_SEAPLANT_PROT=0; HEIX9_FATTYACID=0; HEIX10_SODIUM=0; HEIX11_REFINEDGRAIN=0; HEIX12_ADDEDSUGARS=0; 
#  HEIX13_SATFATS=0;
```

# Grouping stuff
Need to figure out the naming convention 

```{r}
#proc means data=hei0409togroup noprint;
#  by cpartid;
#  var rikcal hei_totveg hei_greensbeans hei_totfruit hei_wholefruit hei_wholegrains hei_dairy hei_totproteins hei_seafoodplantprot 
#		ripfa rimfa risfa hei_sodium hei_refinedgrains hei_addedsugars ripctsfa;
#  output out=dailyhei0409 sum=;
#run;
group_data <- group_by(.data = data, data$Participant.ID)
group_data
```


# Creating components
ref: Development and evaluation of a method for calculating the Healthy Eating Index-2005 using the Nutrition Data System for Research
```{r}
#TOTAL VEGETABLES COMPONENT
data$hei_totveg<-(data$VEG0100+data$VEG0200+data$VEG0300+data$VEG0400+data$VEG0800+data$VEG0450+data$VEG0700+
		data$VEG0600+data$VEG0900+data$VEG0500)/2

#GREENS AND BEANS COMPONENT
data$hei_greensbeans<-(data$VEG0100+data$VEG0700)/2
		
#TOTAL FRUITS COMPONENT
data$hei_totfruit<-(data$FRU0100+data$FRU0200+data$FRU0300+data$FRU0400+data$FRU0500+data$FRU0600+data$FRU0700)/2
  
# WHOLE FRUITS COMPONENT
data$hei_wholefruit<-(data$FRU0300+data$FRU0400+data$FRU0500+data$FRU0600+data$FRU0700)/2
		
# WHOLE GRAINS COMPONENT
data$hei_wholegrains<-data$Whole.Grains..ounce.equivalents.

# DAIRY COMPONENT# INCLUDE DAIRY GROUPS ADDED IN 2019
## missing DML0500, DYF0300 , DYR0300, DYL0300, DOT0900
data$hei_dairy<-(data$DMF0100+data$DMR0100+data$DML0100+data$DMN0100+data$DMF0200+data$DMR0200+
		data$DML0200+data$DML0300+data$DML0400+data$DCF0100+data$DCR0100+data$DCL0100+data$DCN0100+
		data$DYF0100+data$DYR0100+data$DYL0100+data$DYF0200+data$DYR0200+data$DYL0200+data$DYN0100+
		((data$DOT0100)/3)+data$DOT0300+data$DOT0400+data$DOT0500+data$DOT0600)
		
# TOTAL PROTEIN FOODS COMPONENT
data$hei_totproteins<-(data$MRF0100+data$MRL0100+data$MRF0200+data$MRL0200+
		data$MRF0300+data$MRL0300+data$MRF0400+data$MRL0400+data$MCF0200+data$MCL0200+data$MRF0500+
		data$MPF0100+data$MPL0100+data$MPF0200+data$MFF0100+data$MFL0100+data$MFF0200+data$MSL0100+
		data$MSF0100+data$MCF0100+data$MCL0100+data$MOF0100+data$MOF0200+data$MOF0300+data$MOF0400+
		data$MOF0500+data$MOF0600+data$MOF0700+(data$VEG0700*2))

# SEAFOOD AND PLANT PROTEINS COMPONENT
data$hei_seafoodplantprot<-(data$MFF0100+
		data$MFL0100+data$MFF0200+data$MSL0100+data$MSF0100+data$MOF0500+data$MOF0600+data$MOF0700+
		(data$VEG0700*2))
		
## MODERATION
# SODIUM COMPONENT
data$hei_sodium <- data$Sodium..mg./1000
		
# REFINED GRAINS COMPONENT
data$hei_refinedgrains<-data$Refined.Grains..ounce.equivalents.

# ADDED SUGARS COMPONENT
data$hei_addedsugars<-data$Added.Sugars..by.Total.Sugars...g.*4 
data$ripctsfa = data$X..Calories.from.SFA*data$Energy..kcal.
```

# Notes
## Variables not found
DML0500 
DYF0300 
DYR0300
DYL0300
DOT0900

# ADEQUACY
```{r}
# TOTAL VEGETABLES COMPONENT
data$HEIX1_TOTALVEG<-adeq_check(data$hei_totveg, data$HEIX1_TOTALVEG,1.1)

## GREENS AND BEANS COMPONENT;
data$HEIX2_GREEN_AND_BEAN<-adeq_check(data$hei_greensbeans, data$HEIX2_GREEN_AND_BEAN,0.2)

## TOTAL FRUITS COMPONENT;
data$HEIX3_TOTALFRUIT<-adeq_check(data$hei_totfruit, data$HEIX3_TOTALFRUIT,0.8)

## WHOLE FRUITS COMPONENT
data$HEIX4_WHOLEFRUIT<-adeq_check(data$hei_wholefruit, data$HEIX4_WHOLEFRUIT, 0.4)

## WHOLE GRAINS COMPONENT;
data$HEIX5_WHOLEGRAIN<-adeq_check(data$hei_wholegrains, data$HEIX5_WHOLEGRAIN, 1.5)

## DAIRY COMPONENT;
data$HEIX6_TOTALDAIRY<-adeq_check10(data$hei_dairy, data$HEIX6_TOTALDAIRY, 1.3)	

## TOTAL PROTEIN FOODS COMPONENT;
data$HEIX7_TOTPROT<-adeq_check(data$hei_totproteins, data$HEIX7_TOTPROT, 2.5)

## SEAFOOD AND PLANT PROTEINS COMPONENT;
data$HEIX8_SEAPLANT_PROT<-adeq_check(data$hei_seafoodplantprot, data$HEIX8_SEAPLANT_PROT, 0.8)

## FATTY ACIDS COMPONENT;
data$HEIX9_FATTYACID<-fa_adeq(data$Total.Polyunsaturated.Fatty.Acids..PUFA...g., data$Total.Monounsaturated.Fatty.Acids..MUFA...g., data$Total.Saturated.Fatty.Acids..SFA...g.)
```
# Moderation
## Notes
Check sodium, some are negative 
```{r}
## SODIUM COMPONENT
data$HEIX10_SODIUM<-mod_check(data$hei_sodium, data$HEIX10_SODIUM, 1.1, 2.0)


## REFINED GRAINS COMPONENT
data$HEIX11_REFINEDGRAIN<-mod_check(data$hei_refinedgrains, data$HEIX11_REFINEDGRAIN, 1.8, 4.3)

## ADDED SUGARS COMPONENT
data$HEIX12_ADDEDSUGARS<-mod_check(data$hei_addedsugars, data$HEIX12_ADDEDSUGARS, 6.5, 26)

## SATURATED FAT COMPONENT;
data$HEIX13_SATFATS<-mod_check(data$X..Calories.from.SFA, data$HEIX13_SATFATS, 8, 16)
```

#HEALTHY EATING INDEX 2015
```{r}
data$HEI2015_TOTAL_SCORE <- data$HEIX1_TOTALVEG + data$HEIX2_GREEN_AND_BEAN +data$HEIX3 + data$HEIX4_WHOLEFRUIT +
    data$HEIX5_WHOLEGRAIN + data$HEIX6_TOTALDAIRY + data$HEIX7_TOTPROT + data$HEIX8_SEAPLANT_PROT + data$HEIX9_FATTYACID + 
	data$HEIX10_SODIUM + data$HEIX11_REFINEDGRAIN + data$HEIX12_ADDEDSUGARS + data$HEIX13_SATFATS
data$HEI2015_TOTAL_SCORE
```


